#Two Risk Drivers

Initialize library and import data

```{r}
#Libraries
library("here")
library("mvtnorm")

#import data
load(here("data_raw", "Market.rda"))

#unpack Market data
Rf <- Market$rf 
calls <- Market$calls 
puts <- Market$puts 

rf_mat <- as.matrix(Rf)
rf_mat <- cbind(rf_mat, as.numeric(names(Rf)))
rf_mat <- cbind(rf_mat, rf_mat[, 2]*360)
colnames(rf_mat) <- c("rf", "years", "days") 

calls <- cbind(calls, calls[, "tau"]*250)
colnames(calls) <- c("K","tau", "IV", "tau_days")

puts <- cbind(puts, puts[, "tau"]*250)
colnames(puts) <- c("K","tau", "IV", "tau_days")

#import function
source(here("code", "Two Risk Driver", "OptionPricing.R"))
```

Calculate the daily log returns

```{r}
rets_SP <- 100*diff(log(Market$sp500))
rets_VIX <- 100*diff(log(Market$vix))

rets <- cbind(rets_SP, rets_VIX)
rets <- rets[-1,]
colnames(rets) <- c("SP500", "VIX")

```

Generate 10,000 scenarios for 1 week ahead underlying price and the one week ahead VIX value using the normal distribution fitted to the past drivers

```{r}
#Number of days ahead
n_day <- 5

#Assume that the distribution is normally fitted
mu <- apply(rets, 2, mean)

Sigma <- cov(rets)

n_sim <- 10000

#memory allocation
updated_rets <- matrix(NA, n_sim, 2)
rets_day <- matrix(NA, n_sim)

#iterations
set.seed(1234)
for(i in 1:n_day){
    updated_rets <- rmvnorm(mean = mu, sigma = Sigma, n = n_sim)
    
    if(i == 1){
      rets_day <- updated_rets
    }
    else{
      rets_day <- cbind(rets_day,updated_rets)
    }
    
}
head(rets_day)

colnames(rets_day) <- c("Day 1", "Day 1",
                        "Day 2", "Day 2",
                        "Day 3", "Day 3",
                        "Day 4", "Day 4",
                        "Day 5", "Day 5")


daily_rets <- list(SP500 = rets_day[,c(1, 3, 5, 7, 9)],
                   VIX = rets_day[, c(2, 4, 6, 8, 10)])

head(daily_rets$SP500)
```

Now we need to transform the daily log returns to the value of the underlying

```{r}
#Defining an initial value

init <- c(Market$sp500[length(Market$sp500)],
              Market$vix[length(Market$vix)])

#Converting the returns to price
f_convert_log <- function(initial, l_rets){
  #memory allocation
  out <- matrix(NA, nrow(l_rets), ncol(l_rets))
  
  #change the returns to price multiples
  changes <- exp(l_rets/100)

  #initialize first day
  out[, 1] <- as.vector(initial)*changes[,1]
  
  #loop through the matrix 
  for(i in 2:ncol(l_rets)){
     out[,i] <- out[,(i-1)]*changes[,i]
  }
  
  return(out)
}

```

Assign the underlying price and the VIX

```{r}
prices <- list(SP500 = f_convert_log(initial = init[1], daily_rets$SP500),
               VIX = f_convert_log(initial = init[2], daily_rets$VIX))
```

P&L of book of options

```{r, warning = FALSE}
#  maturities for each of the options
T1 <- 20
T2 <- 40
T3 <- 40

# Strikes for the options 
K1 <- 1600 
K2 <- 1605 
K3 <- 1800

#memory allocation
option_price <- lapply(rep(1, 3), # generate three empty matrices of compatible sizes
                  function(x){
                    matrix(NA, 
                           nrow(prices$SP500), 
                           ncol(prices$SP500),
                           dimnames=list(seq(1:nrow(prices$SP500)),
                                         c("T+1", "T+2", "T+3", "T+4", "T+5")
                                         )
                           )
                    }
                  )
names(option_price) <- c("PL1", "PL2", "PL3")

#loop through and find the PL
for(i in 1:n_day){
  #Underlying price at day i
  price_t <- prices$SP500[,i]
  #Underlying volatility at day i
  vol_t <- prices$VIX[,i]
  
  #Call option pricing K = 1605
  c1_vec <- prc_opt(T1-i, K1, calls, rf_mat, price_t, vol_t)
  option_price$PL1[,i] <- c1_vec
  
  #Call option pricing K = 1605 
  c2_vec <- prc_opt(T2-i, K2, calls, rf_mat, price_t, vol_t)
  option_price$PL2[,i] <- c1_vec
 
   #Call option pricing K = 1800
  c3_vec <- prc_opt(T3-i, K3, calls, rf_mat, price_t, vol_t)
  option_price$PL3[,i] <- c1_vec    
}

option_price$PL1
```

```{r}

#Profit of an option -> PL = max(S-K,0)-price
PL_matrices <- lapply(rep(1, 3), # generate three empty matrices of compatible sizes
                  function(x){
                    matrix(NA, 
                           nrow(prices$SP500), 
                           ncol(prices$SP500),
                           dimnames=list(seq(1:nrow(prices$SP500)),
                                         c("T+1", "T+2", "T+3", "T+4", "T+5")
                                         )
                           )
                    }
                  )
names(PL_matrices) <- c("PL1", "PL2", "PL3")

for(i in 1:n_day){
  #spot price of underlying at day i
  spot <- prices$SP500[,i]
  
  #premium at day i for K1
  premium_K1 <- option_price$PL1[, i]
  PL_matrices$PL1[,i] <- max((spot - K1), 0) - premium_K1
  
  #premium at day i for K2
  premium_K2 <- option_price$PL2[, i]
  PL_matrices$PL2[,i] <- max((spot - K2), 0) - premium_K1
  
  #premium at day i for K3
  premium_K3 <- option_price$PL3[, i]
  PL_matrices$PL3[,i] <- max((spot - K3), 0) - premium_K1
  
  
}

```

Graph the PL to find the distribution

```{r}
hist(PL_matrices$PL1)
hist(PL_matrices$PL2)
hist(PL_matrices$PL3)

```




